---
title: "OBIS datasets in GBIF"
output: html_document
editor_options: 
  chunk_output_type: console
---

This notebook addresses the following questions:

- How many OBIS datasets are published to GBIF?
- How many OBIS datasets in GBIF are linked to the OBIS network (GBIF)?
- How many OBIS network (GBIF) datasets are orphaned (no longer published by OBIS)?
- How many OBIS datasets in GBIF do not a citation?

## Data preparation

Load packages:

```{r message = FALSE}
library(robis)
library(rgbif)
library(dplyr)
library(glue)
library(jsonlite)
```

Fetch all OBIS datasets:

```{r message = FALSE}
obis_datasets <- dataset()
```

Fetch all GBIF datasets:

```{r message = FALSE, warning = FALSE}
if (!file.exists("gbif_datasets_endpoints.dat")) {
  
  batch_size <- 100
  gbif_datasets <- data.frame()
  offset <- 0
  
  while (TRUE) {
    url <- glue("https://api.gbif.org/v1/dataset?limit={batch_size}&offset={offset}")
    message(url)
    gbif_subset <- fromJSON(url)$results
    if (length(gbif_subset) == 0) {
      break
    }
    if (!("homepage") %in% names(gbif_subset)) {
      gbif_subset$homepage <- NA
    }
    gbif_datasets <- bind_rows(gbif_datasets, gbif_subset %>% select(key, installationKey, publishingOrganizationKey, title, created, modified, homepage, endpoints))
    offset <- offset + batch_size
  }
  
  # extract archive URLs from GBIF endpoints
  
  find_gbif_archive <- function(endpoints) {
    archives <- endpoints %>% filter(type == "DWC_ARCHIVE")
    if (nrow(archives) > 0) {
      return(archives$url[1])
    } else {
      return(NA)
    }
  }
  
  gbif_datasets_endpoints <- gbif_datasets %>%
    rowwise() %>% 
    mutate(gbif_archive = find_gbif_archive(endpoints))
  
  # save to file
  
  save(gbif_datasets_endpoints, file = "gbif_datasets_endpoints.dat")
  
} else {
  load("gbif_datasets_endpoints.dat")
}
```

Match OBIS and GBIF datasets:

```{r message = FALSE, warning = FALSE}
obis_datasets_merged <- obis_datasets %>%
  left_join(gbif_datasets_endpoints, by = c("archive" = "gbif_archive"))
```

Fetch all OBIS network datasets from GBIF:

```{r message = FALSE}
obis_network <- fromJSON("https://api.gbif.org/v1/network/2b7c7b4f-4d4f-40d3-94de-c28b6fa054a6/constituents?limit=1000")$results
```

## How many OBIS datasets are published to GBIF?

```{r results = "asis"}
obis_in_gbif <- obis_datasets_merged %>%
  filter(!is.na(key))
num_obis_in_gbif <- nrow(obis_in_gbif)
cat(glue("There are {num_obis_in_gbif} active OBIS datasets in GBIF."))
```

## How many active OBIS datasets in GBIF are linked to the OBIS network (GBIF)?

```{r results = "asis"}
obis_in_network <- obis_in_gbif %>%
  filter(key %in% obis_network$key)
num_obis_in_network <- nrow(obis_in_network)
cat(glue("Out of {num_obis_in_gbif} active OBIS datasets in GBIF, {num_obis_in_network} are linked to the OBIS network."))
```

## How many OBIS network (GBIF) datasets are orphaned (no longer published by OBIS)?

```{r results = "asis"}
num_orphaned <- nrow(obis_network) - num_obis_in_network
cat(glue("{num_orphaned} datasets linked to the OBIS network are no longer published by OBIS."))
```

## How many OBIS datasets in GBIF do not a citation?

```{r results = "asis"}
num_no_citation <- sum(is.na(obis_in_gbif$citation))
cat(glue("{num_no_citation} OBIS datasets in GBIF do not have a citation."))
```

## Output

OBIS datasets in GBIF not linked to the network:

```{r message = FALSE}
obis_not_in_network <- obis_in_gbif %>%
  filter(!(key %in% obis_network$key)) %>%
  select(id, url, key)
write.csv(obis_not_in_network, file = "obis_not_in_network.csv", row.names = FALSE)
```
